### 2022-10-18 code/Round 4/Gene expression.Rmd

---
title: "Gene expression"
author: "Kaare D. Tranaes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(broom)
library(ggpubr)
library(AICcmodavg)
library(visdat)
library(rstatix)
library(gt)
library(gtsummary)
library(patchwork)
```

#### Gene expression

barcharts
```{r}
delta_gene_expr <- read_excel("../Data/CleanData/gene_expression_fold_changes.xlsx")

bar_charts_all_genes <- delta_gene_expr %>% 
  mutate(
    groups = recode_factor(  #Change the order of the groups to desired plot order
      groups,
      "con-bf"= "CON-BF",
      "fvt-form" = "VANCO-FF (FVT)",
      "sm-form" = "VANCO-FF (SM)"
      )) %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_col() +
  facet_wrap(~genes,
             ncol = 5, 
             scales = "free") +
  labs(title = "Level of gene expression in ileum tissue",
       y = "Gene expression relative fold change",
       x = "",
       fill = "Groups",
       caption = "Matrinem round 4") +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="lightgrey"), # Change legend box background color
    plot.caption = element_text(color = "darkgrey", size=5)) + 
  scale_fill_brewer(palette="Set1")

bar_charts_all_genes

bar_charts_selected_genes <- delta_gene_expr %>% 
  filter(genes %in% c("TNFa","IL1b", "REG3G")) %>% 
  mutate(
    groups = recode_factor(  #Change the order of the groups to desired plot order
      groups,
      "con-bf"= "CON-BF",
      "fvt-form" = "VANCO-FF (FVT)",
      "sm-form" = "VANCO-FF (SM)"
      )) %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_col() +
  facet_wrap(~genes,
             ncol = 5, 
             scales = "free") +
  labs(title = "Level of gene expression in ileum tissue - significant changes",
       y = "Gene expression relative fold change",
       x = "",
       fill = "Groups",
       caption = "Matrinem round 4") +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="lightgrey"), # Change legend box background color
    plot.caption = element_text(color = "darkgrey", size=5)) + 
  scale_fill_brewer(palette="Set1")

bar_charts_selected_genes
```