### 2022-10-18 code/cleaning_raw_data NEW.Rmd

---
title: "document_title"
author: "Kaare D. Tranaes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
```

# Raw data cleaning document

::: {#Project Description}
This is a documentation of generating the working files from the raw data files. Every step of the process is explained below.  
The raw data files can be accessed in the RawData folder associated with this project.  
Once cleaned up, a new data file for further analysis is created and stored in CleanData folder.
:::

```{r load required libraries}
library(readxl) 
library(tidyverse)
library(writexl)
library(broom)
```
## Weight development data
```{r load data}
import_round1 <- read_excel("../Data/RawData/bodyweight_round_1.xlsx"
        #                     , col_types = c("text","text", "text", "text", 
        # "text", "text", "text", "text", "numeric", 
        # "numeric", "numeric", "text")
        ) %>% 
  select(-`Pup ID`) %>% 
  rename(sample_id = `Sample ID`,
         exp_number = Round,
         received_antibiotics = `Treatment with AB`, 
         maternal_treatment = `Type of AB`,
         type_of_feed = `Feeding`,
         group = Group,
         cage = Mother,
         bodyweight_baseline = `Baseline weight`,
         bodyweight_baseline_plus_24h = `Middle of study weight`,
         bodyweight_baseline_plus_48h = `Weight End`)

import_round2 <- read_excel("../Data/RawData/bodyweight_round_2.xlsx"
        #                     ,col_types = c("text", "text", "text", 
        # "text", "text", "text", "text", "numeric", 
        # "numeric", "numeric", "numeric","numeric", "numeric")
        ) %>% 
  rename(sample_id = `ID`,
         exp_number = Round,
         received_antibiotics = `Treatment with AntiBiotic`, 
         maternal_treatment = `Type of AB`,
         type_of_feed = `Feed`,
         group = Groups,
         cage = Cage,
         bodyweight_birthday_group_average = birthweight,
         bodyweight_birthday_plus_24H_group_average = `birth+1day`,
         bodyweight_birthday_plus_48H_group_average = `birth+2days (group mean)`,
         bodyweight_baseline = `Baseline`,
         bodyweight_baseline_plus_24h = `Study day 1`,
         bodyweight_baseline_plus_48h = `End of study`)

import_round3 <- read_excel("../Data/RawData/bodyweight_round_3.xlsx"
        ) %>% 
  select(-id) %>% 
  rename(sample_id = `Sample ID`,
         exp_number = Exp_number,
         received_antibiotics = `Antibiotic yes/no`, 
         maternal_treatment = `Type of AB`,
         type_of_feed = `Type of feeding`,
         group = Group,
         bodyweight_baseline = `Start`,
         bodyweight_baseline_plus_24h = `Day 2`,
         bodyweight_baseline_plus_48h = `End weiight`) %>% 
  mutate(sample_id = as.numeric(sample_id),
          exp_number = as.numeric(exp_number))

import_round4 <- read_excel("../Data/RawData/bodyweight_round_4.xlsx"
        #                     ,col_types = c("text", "text", "text", 
        # "text", "text", "text", "text", "text", 
        # "numeric", "numeric", "numeric")
        ) %>% 
  select(-mark) %>% 
  rename(sample_id = `Sample ID`, 
         group = group,
         bodyweight_baseline = `weight 0H`,
         bodyweight_baseline_plus_24h = `Weight 24H`,
         bodyweight_baseline_plus_48h = `Weight 48H`) %>% 
  mutate(cage = as.numeric(cage))

weights_all_rounds <- bind_rows(import_round1, import_round2, import_round3, import_round4) %>% 
  relocate(comments, .after = bodyweight_birthday_plus_48H_group_average) %>% 
  mutate(received_antibiotics = received_antibiotics %>% trimws(),
         maternal_treatment = maternal_treatment %>% trimws()) %>% 
  mutate(type_of_feed = if_else(type_of_feed == "Formula", "FF",
                                      if_else(type_of_feed == "Breastmilk", "BF",
                                              if_else(type_of_feed == "Breastfeeding", "BF", type_of_feed)))) %>% 
  #mutate_if(is.character,tolower) %>% 
  mutate(maternal_treatment = if_else(maternal_treatment == "Vanco", "VANCO",
                                      if_else(maternal_treatment == "Vancomycin", "VANCO",
                                      if_else(maternal_treatment == "Genta", "GENTA",
                                              if_else(maternal_treatment == "Gentamycin", "GENTA",
                                                      if_else(maternal_treatment == "Water", "CON",
                                                              if_else(maternal_treatment == "Mix", "AB",
                                                              maternal_treatment))))))) %>% 
  unite(group2, maternal_treatment:type_of_feed, sep= "-", remove = FALSE)

View(weights_all_rounds)
rio::export(weights_all_rounds, file = "../Data/CleanData/weight_development.xlsx", overwrite = TRUE)
```

## Study IDs 
```{r}
study_chars <- weights_all_rounds %>% 
  select(1:7)

rio::export(study_chars, file = "../Data/CleanData/study_id_info.xlsx", overwrite = T)

```

## Cytokines from mesoscale assays
```{r}
study_chars <- read_excel("../Data/CleanData/study_id_info.xlsx")

mesoscale_round_1 <- read_excel("../Data/RawData/cytokines_round_1_r.xlsm")

mesoscale_round_2_and_3 <- read_excel("../Data/RawData/cytokines_round_2_and_3.xlsx") %>% 
  rename("tissue_weight_mg" = "tissue_weight_(mg)", "exp_number" = "exp_no")

mesoscale_round_4 <- read_excel("../Data/RawData/cytokines_round_4.xlsx") %>%
  rename("group" = "Group", "exp_number" = "exp_no")

mesoscale_all_rounds <- bind_rows(mesoscale_round_1, mesoscale_round_2_and_3, mesoscale_round_4) %>% 
  select(-group, -treatment, - exp_number)

cytokines <- left_join(study_chars, mesoscale_all_rounds, by = c("sample_id" = "sample")) %>% 
  relocate(tissue:tissue_weight_mg, .after= type_of_feed)

#rio::export(cytokines, file = "../Data/CleanData/cytokines.xlsx")
```

## Intestinal permeability - FITC-Dextran assays
```{r}
fitc_round_2_and_3 <- read_excel("../Data/RawData/FITC_round_2-3.xlsx") %>% 
  rename(exp_number = exp_no) %>% 
  mutate("result_µg/ml_undiluted" = `result_µg/ml`*dilution_factor)

fitc_round_4 <- read_excel("../Data/RawData/FITC_round_4.xlsx") %>% 
  mutate("result_µg/ml_undiluted" = `result_µg/ml`*dilution_factor)

fitc_all_rounds <- bind_rows(fitc_round_2_and_3, fitc_round_4) %>% 
  select(-group, - exp_number, -cage)

fitc <- left_join(study_chars, fitc_all_rounds, by = c("sample_id" = "sample")) %>%  
  filter(exp_number != "1")  # FITC-Dextran not performed in experimental round 1.

rio::export(fitc, file = "../Data/CleanData/fitc.xlsx")
```

## Gene expression - qPCR results
```{r}
### Start here. Or do it in excel.
study_chars <- read_excel("../Data/CleanData/study_id_info.xlsx")
qpcr <- read_excel("../Data/RawData/all_genes_ct_values.xlsx")

qpcr <- qpcr %>% 
  mutate(mean_cqt = (Cq1 + Cq2)/2) %>% 
  pivot_wider(id_cols = Sample, 
              names_from = Gene,
              values_from = mean_cqt)

gene_expr <- study_chars %>%   #### Maybe export this to excel and try to create double delta ct analysis there. 
  filter(sample_id %in% 89:112) %>% 
  left_join(., qpcr, by = c("sample_id" = "Sample"))

delta_gene_expr <- gene_expr %>%
  rename(ct_value_house_gene_HRPT = `House gene (HRPT)`) %>% 
  pivot_longer(cols = CD8:VEGFa,
               names_to = "genes",
               values_to = "ct_value") %>%
  select(sample_id, groups, genes, ct_value, ct_value_house_gene_HRPT) %>% 
    filter(!is.na(ct_value)) %>%
  mutate(dct = ct_value-ct_value_house_gene_HRPT) %>% 
  group_by(genes, groups) %>% 
  mutate(avg_group_dct_val = mean(dct)) %>% 
  group_by(genes) %>% 
  mutate(ddct = dct - (avg_group_dct_val[groups=="con-bf"])) %>% 
  mutate("2^-(ddct)" = 2^-ddct) %>% 
  group_by(genes, groups) %>% 
  mutate(avg_group_2_exp_ddct_val = mean(`2^-(ddct)`)) %>%
  group_by(genes) %>% 
  mutate(relative_fold_change = `2^-(ddct)`/(avg_group_2_exp_ddct_val[groups=="con-bf"]))

# delta_gene_expr %>% 
#   group_by(genes, groups) %>% 
#   summarise(mean_fold_change = mean(relative_fold_change))

#rio::export(delta_gene_expr, file = "../Data/CleanData/gene_expression_fold_changes.xlsx")

```

```{r}

bar_charts <- delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_col() +
  facet_wrap(~genes,
             ncol = 5) +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )
```

```{r}
delta_gene_expr %>%
  filter(genes %in% "CD8") %>% 
  anova(relative_fold_change ~ groups, data = .)
```