### 2022-10-18 code/lin_reg_vanco_interaction.Rmd

---
title: "Untitled"
author: "Kaare D. Tranaes"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
library(readxl)
library(tidyverse)
library(broom)
library(jtools)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#_______Read in data______________________________________
```{r}
fitc <- read_excel("../Data/CleanData/fitc.xlsx") %>% 
  rename(values = `result_µg/ml_undiluted`) %>% 
  drop_na(values)
```
#_________________

#_______Test effect of antibiotics and formula feeding_______ 
```{r}
fitc %>% 
  filter(exp_number == 3) %>% 
  lm(values ~ received_antibiotics + type_of_feed, data = .) %>% summary()  
```
#Turns out not to be significant: Multiple R-squared:  0.09071, p-value: 0.1894
#______________

#_______Test effect of interaction between antibiotics and formula feeding_______
```{r}
fitc %>% 
  filter(exp_number == 3) %>% 
  lm(values ~ received_antibiotics*type_of_feed, data = .) %>% summary()  
```
#Turns out not to be significant: Multiple R-squared:  0.1244, p-value: 0.205
#______________

#_______Test effect of type of antibiotics________________
```{r }
lm_vanco <- fitc %>% 
  filter(exp_number == 3) %>% 
  lm(values ~ maternal_treatment*type_of_feed, data = .)
```
#Significant effect of VANCO-treatment: "maternal_treatmentVANCO:type_of_feedFF   Estimate(58.538)     SD(33.151)   (tvalu)1.766    P =0.087  
#Multiple R-squared:  0.1244, p-value: 0.008879
#______________

```{r results = 'asis'}
export_summs(lm_vanco)
```