### 2022-10-20 submission/manuscript.Rmd
Progressed on manuscript.

```
output: html_document

```

## Introduction

- NEC is caused by bacterial dysbiosis - we think. 
- (AND) high-osmolar formula feed is hard on intestinal lining. 
- (AND) Chen et al. saw NEC-like symptoms with MAT.
- (BUT) No enough on its own plus too old mice. 
- (Therefore) We want to see if MAT + FF will result in more severe NEC-like symptoms. (AIM)
- (AND) FvT are currently being studied as potential remidies to NEC - thought to be mediated
via stabilizing effect on microbiome. 
- (Therefore) We want to see if A) we manage to induce NEC, are we able to "cure" it, or B) in 
case of no NEC, is FvT on it's own safe or associated with any adverse effects?

#### Rationale:

*Chen et al.*, showed that Maternal Antibiotics Treatment (**MAT**) resulted in mild NEC-like symptoms of mice offspring. We believed that MAT combined with formula feeding (**FF**) of offspring would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

## Results (decide on which voice to use, active or passive. Read other papers.)

Experiment no. 4 was the goal, but before we could test our overall hypothesis we needed to find out which combination of antibiotics and feeding scheme would results in the most NEC-like phenotype.
These research questions were investigated in 3 pre-experiments (PreExp 1-3), each with their own research questions.

#### Model feasibility
- Our first experiement (PreExp 1) was designed to test wheteher  it was possible to keep mice separated from the mother alive from postnatal day 3 and 48 hours onwards. 
- `r nrow(preexp_1)` mice were born to 6 different mothers. Of these 32 survived and were included in the analysis. After fine-tuning the feeding method we we able to successfully feed all surviving (32) mice every 3 hours for 2 days (48 hours).
(Conclusion) - Visual inspection of intestines during sampling concluded that we had not created a NEC phenotype in any treatment group.
(BUT) - We found that AB-FF pups consistently looked different that CON-BF in the level of cytokine expression in the ileum, with statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES. 
- A multiple linear regression analysis, evaluating the effects of MAT (AB or CON) and type of feed (BF of FF) separately and interactively found statistically significant effects of formula feeding for REG_MODEL_CYTOKINES$TERM == TYPEOFFEED, with the greatest effect estimate of ESTIMATE$IL-6 , P-val$IL-6.

We therefore decided to test different types of antibiotics.

#### Microbiome dysbiosis
Based on the results from Preexp 1 we knew that we could perform the experimental setup without too many casualties. 
- We had seen that formula-feeding did affect the expression of cytokines, but visual inspection of the intestines showed that we had not managed to create a NEC-like phenotype. 
- After considering the results we realized that our assumptions were based on a dysbiotics microbiome in the mother leading to less immunizing milk. However, we had likely not created a dysbiosis as much as an overall depletion by administering a broad-spectered antibiotic. 
We therefore hypothezised that targeting specific stains of bacteria would lead to a more dysbiotic microbiome and allow for greater disbiosis in the offspring.

### Round 4
- We treated XX mothers with Vancomycin (1 g/l) from pregnancy-day 15 until delivery, where vancomycin-mixed drinking water was replaced with regular water.
- 2 days after birth xx of the 26 pups were separated from the mother and received XX µl FvT solution or an equal amount of SM buffer (control treatment). The remaning xx of 26 pups remained with the mother for breastfeeding thoughout the study.
- Pups separated from their mother were formula fed (FF) with a solution of XXXXX milk formulation (see materials) every 3 hours for 48 hours. (futher setup description see materials and methods).
- All 26 pups survived the 48 hour study. On the final day of the study pups fasted for XX hours before given xx µl FITC Dextran (as per materials protocol). 
- 2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration. Intestinal samples were collected for gene expression assays, histological assessment, pro-inflmmatory cytokine analysis and gut microbiome sequencing.

* Gut permeability: 
- We found the lowest concentration of serum FITC Dextran in the group treated with FF and Fvt (XXXX Xg/mL). This was lower than even the breastfed controls, which had the highest concentration of the three groups (**FIGURE**).

* Cytokine expression
- High levels of XX cytokine was foudn in group XX.

* Histological analysis
- We found no significant changes in the mucin layer between the treatment group (Fvt) Fvt-Control (SM) or formula-feed control (VANCO-BF). The thickness of the stain was the same for all groups.
* This supports the impression we got from the visual inspection of the intestines during sampling.

* Gene expression 
We standardized the gene expression to the control group and the house gene. However, we found statistically significant difference in expression of the house-gene, making the delta-delta measure bad. Comparing (on what other basis) in stead.

* Gut microbiome
Differneces?

## Discussion

## Materials and methods
Describe Pre-Exp 1-3 and knowledge gained from them.