### 2022-10-27 submission/manuscript.Rmd

---
title: "Results section"
author: "Kaare"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Code

# Manuscript

## Introduction

- NEC is caused by bacterial dysbiosis - we think. 
- (AND) high-osmolar formula feed is hard on intestinal lining. 
- (AND) Chen et al. saw NEC-like symptoms with MAT.
- (BUT) No enough on its own plus too old mice. 
- (Therefore) We want to see if MAT + FF will result in more severe NEC-like symptoms. (AIM)
- (AND) FvT are currently being studied as potential remidies to NEC - thought to be mediated
via stabilizing effect on microbiome. 
- (Therefore) We want to see if A) we manage to induce NEC, are we able to "cure" it, or B) in 
case of no NEC, is FvT on it's own safe or associated with any adverse effects?

## Background

- Why we think MAT leads to intestinal dysbiosis?
    - Less immunizing milk?
-

## Results

The main goal of this thesis project was to create a NEC mouse model by combining Maternal Antibiotics Treatment with Formula-Feeding. This hypothesis was tested in a series of experiments (Exp 1-4), each successive experiment building on the results of the former. 
The following is a description of each experiment. An abridged version of each experiment is provided, followed by a detailed  description of experimental setup and results.

### Exp 1: Model feasibility

#### Rationale:
*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim: 
- To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
- Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
- Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visaul inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 1. **Date of exp** 
*Antimals and experimental setup* See "Animal experimental setup" in methods.
- The experimental setup has been described in detail in the methods seciton. 
- Detials specific for Exp-1 are outlined below. 
- Briefly, six male and 18 female C57BL/6N mice were co housed for 24 hours in 6 cages (1 male + 3 female each).
- Two weeks after mating, pregnant females were separated evenly into maternal antibiotic treatment (MAT) and water-only controls (CON). 
- Antibiotics (Gentamicin, Vancomycin, Neomycin, and Ampicillin (all 0.5 g/L)) was mixed into the drinking water of the MAT mice while controls received water only.
- Antibiotic treatment commenced on gestational day 15 and continued until delivery. 
- Each mother was housed with her offspring in separate cages during the first 3 days after birth. 
- On the 4th day, `r nrow(preexp_1)` mice, XXX (Filter `r nrow(preexp_1)` for feed type) from antibiotics treated mothers and XXX (Filter `r nrow(preexp_1)` for feed type) were divided into four experimental groups: MAT + formula-feeding (AB-FF) MAT + breastfeeding (AB-BF), Water + formula-feeding (CON-FF), and CON + breastfeeding (CON-BF). 
- Formula-feeding was performed as per METHODS AND MATERIALs. In Exp-1 we administered  0.08 ml formula for pups >2g, and 0.06 for pups ≤ 2g every 3 hours throughout the study. In later experiments this dose was changed, and feeding intervals were inreased to every 4 hours from midnight til 08 AM.

- `r nrow(preexp_1_survived)` of `r nrow(preexp_1)` pups survived until sampling.
- Those that did not survive were euthanize because of esophageal perforation during feeding. After fine-tuning of the feeding methods no pups were eithanized as a result of reaching humane endpoints, and the pups generally were not very sick. 
- While significantly less than their breastfed controls, all pups increased their heavyweight during the experiment, attesting to their generally reasonable health status.
- The pups were physically active and reacted to changes in light and to physical sitmuli.
- Skin color remained bright purple and fur development was visible thoughout the experiment. 
- No signs of intestinal inflammation were evident from the visual inspection of intestines during sampling. Our subjective observation of the intestines were small degrees of bloating in formula-fed pups.
- Analysis of pro-inflammatory cytokines showed statistically significant changes in the expression of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Discussion of experimental results - Exp 1.
- Directionality of change varied, with alternating over- and under-expression of various cytokines. As can be seen in [*Cytokine plot of signif results *] pups tended to group based on feeding-scheme, with both formula-fed groups being either suppressed og over-expressed together, compared with breastfed groups, irrespective of antibiotics treatment. 
- This observation is supported by a a multiple linear regression analysis, evaluating the effects of MAT (AB or CON) and type of feed (BF of FF) separately and interactively. The analysis found statistically significant effects of formula feeding for REG_MODEL_CYTOKINES$TERM == TYPEOFFEED, with the greatest effect estimate of ESTIMATE$IL-6 , P-val$IL-6.

- In summary, formula-feeding seemed to have greater effect on the inflammatory state of the pups than whether they were weened by antibiotics treated mothers. 
- This is surprising given the findings of Chen et al., who used an identical protocol to our AB-BF vs. CON-BF (Maternal antibiotics treatment and weening with the mother for 7 days). They claim that MAT during pregnancy resulted in significant alterations in the intestinal microbiota. lower body weight at birth and on day 7, visible intestinal injury and inflammation and disruption of the intestinal barrier. 
- Our initial analysis of weight and cytokine expression did not indicate a similar effect of MAT (evaluations of gut permeability, intestinal gene expression, histological analysis and gut microbial composition were conducted in later experiments). 
- One possible explanation is an effect of dosage. While Chen et al., used a concentration of 1 g/l, we went with 0.5 g/l, based on experience from previous studies.
- We therefore decided to increase the dosage to 1 g/l for subsequent experiments,
and to include further measures of intestinal dysfunction.
- We also decided to measure the birth weight of all pups, in addition to body weight at the beginning of the experiment (day 3) and end (day 5).

### Experiment 2: Effects of various types of antibiotics - (Microbiome dysbiosis)

#### Rationale:
- Exp 1 demonstrated that the type of feed (breastmilk or formula) had a greater effect on cytokine expression than maternal antibiotics treatment. 
- Central to our hypothesis of NEC development is an intestinal microbiome dysibiosis, and a broad-spectrered antibiotic treatment potentially resulted in an overall depletion rather than the manifisation of a dysbiosis. 
- We therefore hypothesized that targeting specific strains of bacteria in the mother would lead to a more dysbiotic microbiome and allow for greater subsequent dysbiosis in the offspring.

#### Aim: 
- To test the effect of gram-positive-specific antibiotics versus gram-negative-specific antibiotics, without formula feeding on intestinal permeability and cytokine expression. 
- To see whether a larger dose of antibiotics (1 g/l) causes decreased body weight at birth and in day 5 compared with non-treated controls.

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visual inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 2.

*Antimals and experimental setup*
- The breeding setup in Exp 2 was identical to Exp 1. 
- On gestational day 15 vancomycin, gentamicin or control (tab water) was added to the drinking water pregnant females, forming 3 experimental groups; VANCO, GENTA and CON.

*Weight development*
- We checked the cages twice daily from gestational day 18 and measured the weight of newly-born pups. 
- VANCO and GENTA mothers gave birth on the 25th of february (getational day 20) with mean offspring body weights of `r mean(bodyweight_birth_vanco` (N = `r nrow(bodyweight_birth_vanco`) and `r mean(bodyweight_birth_genta` (N = `r nrow(bodyweight_birth_vanco`) respectively. 
- The CON mother gave birth on the 26th of february (gestational day 21) with mean offspring body weight of `r mean(bodyweight_birth_con` (N = `r nrow(bodyweight_birth_vanco`). 
- This difference in body weight at birth was statistically significant (p = `r lm_weight_exp2$p.value`), but adjusting for length of gestation erased the statistical significance (p = `r lm_weight_exp2_adj$p.value`). 
- On day 3 and 5 this difference was in body weight was no longer detectable. *(Figure of body weights in exp 2)*

*Intestinal permeability - FITC-Dextran administration* 
- On the day of euthanasia pups were separated from the mothers at 07:00 (halting breastfeeding). 
- At 10:00, 50 µL FITC-dextran (0.045 mg/µL) was administered orally using a PICC line and a 1 mL syringe for a final dose of 750 mg/kg. 
- Euthanasia and sample collection commenced at noon.
- Pups were euthanized by cervical dislocation and blood samples were collected and spun in Eppendorf tubes. 
- FITC molecules were detectable in all groups. The highest and lowest concentration was found in the GENTA (`r fitc_exp2_genta` µg/ml) and VANCO (`r fitc_exp2_vanco` µg/ml) groups, respectively, with the CON group in between (`r fitc_exp2_vanco` µg/ml). This difference was statistically significant (p = `r anova_fitc_exp2`, one-way analysis of variance).

*Cytokine expression*
- Tissue weight-adjusted results of cytokine concentration showed significant differences between groups for `r signif_cyto_exp2`. 
- A statistically significant difference in the average weight of the intestinal biopsy was found ( p = `r tissue_weight_exp2$p.value`), with XXXX having the heaviest (`r tissue_weight_exp2_`) sample weights and XXXX having the lowest.
- Un-adjusted for tissue weight there were no significant differences between the treatment groups.

#### Discussion of experimental results - Exp 2.
- Directionality of change varied, with alternating over- and under-expression of various cytokines. As can be seen in [*Cytokine plot of signif results *] pups tended to group based on feeding-scheme, with both formula-fed groups being either suppressed og over-expressed together, compared with breastfed groups, irrespective of antibiotics treatment. 
- This observation is supported by a a multiple linear regression analysis, evaluating the effects of MAT (AB or CON) and type of feed (BF of FF) separately and interactively. The analysis found statistically significant effects of formula feeding for REG_MODEL_CYTOKINES$TERM == TYPEOFFEED, with the greatest effect estimate of ESTIMATE$IL-6 , P-val$IL-6.

- In summary, formula-feeding seemed to have greater effect on the inflammatory state of the pups than whether they were weened by antibiotics treated mothers. 
- This is surprising given the findings of Chen et al., who used an identical protocol to our AB-BF vs. CON-BF (Maternal antibiotics treatment and weening with the mother for 7 days). They claim that MAT during pregnancy resulted in significant alterations in the intestinal microbiota. lower body weight at birth and on day 7, visible intestinal injury and inflammation and disruption of the intestinal barrier. 
- Our initial analysis of weight and cytokine expression did not indicate a similar effect of MAT (evaluations of gut permeability, intestinal gene expression, histological analysis and gut microbial composition were conducted in later experiments). 
- One possible explanation is an effect of dosage. While Chen et al., used a concentration of 1 g/l, we went with 0.5 g/l, based on experience from previous studies.
- We therefore decided to increase the dosage to 1 g/l for subsequent experiments,
and to include further measures of intestinal dysfunction.
- We also decided to measure the birth weight of all pups, in addition to body weight at the beginning of the experiment (day 3) and end (day 5).

### Experiment 3 - Interaction effects between antibiotics treatment and formula feeding

#### Rationale:
*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim: 
- To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
- Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
- Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visaul inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 3.
- In round 3 we repeated the experimental setup with regards to MAT, adding either Vancomycin, Gentamicin or sterile water to the drinking water of pregnant "females" (use same word throughout(mothers)). We repeated the group-setup from PreExp  1 and  separated pups into breastfeeding or formula feeding, for a total of 6 groups (see overview table). 
- All pups survived the experiment and were included in the final analysis. 
- 2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration. Intestinal samples were collected for gene expression assays, histological assessment, pro-inflammatory cytokine analysis and gut microbiome sequencing.

* Gut permeability: 
- Formula-fed pups consistently showed higher levels of Serum FITC concentration compared to breastfed pups, irrespective of MAT status. VANCO-FF and VANCO-BF had the highest and lowest levels of FITC, respectively. 
- VANCO-BF was the group with the lowest FITC in PreExp 2.
- This is indicative of an interaction-effect with vancomycin and formula feeding. (**FIGURE**).

* Cytokine expression
- We found significant differences in concentrations of XXSignificantSCYTOKINESXX. Linear regression analysis indicated the biggest detminant of differneces in outcomes was formula feeding, with the type of antibiotic being of lesser importance compared. Directionally the trend was the same for all cytokines. 
- IL-2: significantly lower concentrations in all formula-fed pups compared with CON-BF.
- IL-6: Significantly greater levels in all formula-fed pups compared with CON-BF.
- INFg: significantly lower concentrations in all formula-fed pups compared with CON-BF.
- KCGRO: Significantly greater levels in all formula-fed pups compared with CON-BF.

### Experiment 4

#### Rationale:
*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim: 
- To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
- Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
- Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visaul inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 4.
- We treated XX mothers with Vancomycin (1 g/l) from pregnancy-day 15 until delivery, where vancomycin-mixed drinking water was replaced with regular water.
- 2 days after birth XX of the 26 pups were separated from the mother and received XX µl FvT solution or an equal amount of SM buffer (control treatment). The remaning xx of 26 pups remained with the mother for breastfeeding thoughout the study.
- Pups separated from their mother were formula fed (FF) with a solution of XXXXX milk formulation (see materials) every 3 hours for 48 hours. (futher setup description see materials and methods).
- All 26 pups survived the 48 hour study. On the final day of the study pups fasted for XX hours before given xx µl FITC Dextran (as per materials protocol). 
- 2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration. Intestinal samples were collected for gene expression assays, histological assessment, pro-inflammatory cytokine analysis and gut microbiome sequencing.

* Gut permeability: 
- We found the lowest concentration of serum FITC Dextran in the group treated with FF and Fvt (XXXX Xg/mL). This was lower than even the breastfed controls, which had the highest concentration of the three groups (**FIGURE**). 
- The VANCO-FF (Fvt) compared with VANCO-FF from PreExp 3 this change was significant/non-significant. 
- It is reasonable/unreasonable to compare the rounds, given the similarity/lack of similarity in response of the VANCO-BF groups from Exp3(XXX) and NEC-FvT (XXXX)

* Cytokine expression
- Only TNFa was significantly different in Exp 4, with VANCO-FF (FvT) having the highest level (XXX), signifiacntly greater than both VANCO-BF and VANCO-FF (SM) (XXX and XXXX respectively). 
- The VANCO-FF (FvT) had lower/higer absolute levels of cytokine concentration compared to both VANCO-BF and VANCO-FF from Exp3. 
- THis comparison is reasonable/unreasonable given the similarity/lack of similarity in response of the VANCO-BF groups from Exp3(XXX) and VANCO-FF (SM) from Exp 4(XXXX).

### Post experimental analyses and Discussion of results 
* Histological analysis - Mention and Discuss this after finishing the result section. 
- Histological image analysis was performed after all trials had finished. 
- We saw no degradation of mucin layer thickness in any of the pups. This supports the visial inspections of intestines that was made during sampling. 
- We found no significant changes in the mucin layer between the treatment group (Fvt) Fvt-Control (SM) or formula-feed control (VANCO-BF). The thickness of the stain was the same for all groups.

* Gene expression

We standardized the gene expression to the control group and the house gene. However, we found statistically significant difference in expression of the house-gene, making the delta-delta measure bad.

* Gut microbiome

## Discussion