### 2022-11-04 submission/results.Rmd

---
title: ''
output: pdf_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

<!--# Code -->

<!--# Run scripts for plots on another file and save plots on img folder. Then use include_graphics to include plots where relevant -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r loading libraries, message=FALSE, warning=FALSE}
library(readxl)
#library(tidyverse)
library(broom)
library(ggpubr)
#library(AICcmodavg)
#library(visdat)
#library(rstatix)
#library(gt)
#library(gtsummary)
library(patchwork)
#library(jtools)
```

```{r exp1-weight, include=FALSE}
#____Load overview of sample IDs______ 
pups <- read_excel("data/processed/weight_development.xlsx")

#____ Separate into experimental rounds ____#
preexp_1 <- pups %>% 
  filter(exp_number == "1")

exp1_survived <- preexp_1 %>% 
  filter(!is.na(bodyweight_baseline_plus_48h)) %>% 
  nrow()
#____ ____#
```

<!-- Manuscript -->

The main goal of this thesis project was to create a NEC mouse model by combining Maternal Antibiotics Treatment with Formula-Feeding.
This hypothesis was tested in a series of experiments (Exp 1-4), each successive experiment building on the findings from the former.
The following is a description of each experiment.
An abridged version of each experiment is provided, followed by a detailed description of experimental setup and results.

### Exp 1: Model feasibility

#### Rationale:

-   *Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring.
-   XXX *[Find paper]* found formula-feeding to be damaging to the intestine.
-   One of the primary causal factors of NEC is formula-feeding (paper on what is thought to cause NEC).
-   We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim:

-   To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
-   Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
-   Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:

-   Survival rate: `r`exp1_survived`of`r nrow(preexp_1)`(`r (exp1_survived / nrow(preexp_1))\*100\`%)
-   Visual inspection: no clear signs of intestinal inflammation
-   Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 1. **Date of exp**

*Antimals and experimental setup* See "Animal experimental setup" in methods.
- The experimental setup has been described in detail in the methods section.
- Detials specific for Exp-1 are outlined below.
- Briefly, six male and 18 female C57BL/6N mice were co-housed for 24 hours in 6 cages (1 male + 3 female each).
- Two weeks after mating, pregnant females were separated evenly into maternal antibiotic treatment (MAT) and water-only controls (CON).
- Antibiotics (Gentamicin, Vancomycin, Neomycin, and Ampicillin (all 0.5 g/L)) were mixed into the drinking water of the MAT mice while controls received water only.
- Antibiotic treatment commenced on gestational day 15 and continued until delivery.
- Each mother was housed with her offspring in separate cages during the first 2 days after birth.

```{r exp1-groups}
#____Separate into groups ______ 
AB_FF <- nrow(preexp_1 %>% 
  filter(group2 == "AB-FF"))

AB_BF <- nrow(preexp_1 %>% 
  filter(group2 == "AB-BF"))

CON_FF <- nrow(preexp_1 %>% 
  filter(group2 == "CON-FF"))

CON_BF <- nrow(preexp_1 %>% 
  filter(group2 == "CON-BF")) 
#____ ____#

```

-   On day 3, mice from water or antibiotics treated mothers were separated and assigned to either formula feeding (FF) or breastfeeding (BF) forming four experimental groups: AB + formula-feeding (AB-FF, N = `r AB_FF` ) AB + breastfeeding (AB-BF, N = `r AB_BF` ), Water + formula-feeding (CON-FF, N = `r CON_FF` ), and CON + breastfeeding (CON-BF, N = `r CON_BF` ).

-   Formula-feeding was performed as per METHODS AND MATERIALs.
    In Exp-1 we administered 0.08 ml formula for pups \>2g, and 0.06 for pups ≤ 2g every 3 hours throughout the entire study.
    In later experiments this dose was changed, and feeding intervals were increased to every 4 hours between midnight and 08 AM.

-   `r exp1_survived` of `r nrow(preexp_1)` pups survived the treatments.

-   Those that did not survive were euthanize because of esophageal perforation during feeding.

-   After fine-tuning of the feeding methods we were able to feed the pups without further harm.

-   We never had to euthanize a pup that reached our humane endpoints, since all the surviving pups were in generally good shape.

-   While significantly lighter than their breastfed controls, all pups increased their bodyweight during the experiment, attesting to the efficacy of our feeding regiment.

-   The pups were physically active and responded to changes in light and physical stimuli.

-   Skin color remained bright purple and fur development was visible throughout the experiment.

-   No signs of intestinal inflammation were evident from the visual inspection of intestines during sampling.
    Our subjective observation of the intestines were small degrees of bloating in formula-fed pups.

**Cytokines**

`r nrow(cytokines_1)` samples were available for analysis of the proinflammatory cytokines `r names_of_cytokines`.
Concentrations were measured in 1-cm sections of proximal/distal ileum tissue using MSD Mesoscale Kits.
Statistically significant changes in the expression of `r names_of_cytokines`.
- Directionality of change varied, with alternating over- and under-expression of various cytokines.
As can be seen in [*Cytokine plot of signif results* ] pups tended to group based on feeding-scheme, with both formula-fed groups being either suppressed or over-expressed together, compared with breastfed groups, irrespective of antibiotics treatment.

-   This observation is supported by a a multiple linear regression analysis, evaluating the effects of MAT (AB or CON) and type of feed (BF of FF) separately and interactively. The analysis found statistically significant effects of formula feeding for REG_MODEL_CYTOKINES$TERM == TYPEOFFEED, with the greatest effect estimate of ESTIMATE$IL-6 , P-val\$IL-6.

#### Commentary and interpretation of results - Exp 1.

-   In summary, formula-feeding seemed to have greater effect on the inflammatory state of the pups than whether they were weened by antibiotics treated mothers.
-   This is surprising given the findings of Chen et al., who used an identical protocol to our AB-BF vs. CON-BF (Maternal antibiotics treatment and weening with the mother for 7 days).
-   They claim that MAT during pregnancy resulted in significant alterations in the intestinal microbiota. lower body weight at birth and on day 7, visible intestinal injury and inflammation and disruption of the intestinal barrier.
-   Our initial analysis of weight and cytokine expression did not indicate a similar effect of MAT (evaluations of gut permeability, intestinal gene expression, histological analysis and gut microbial composition are conducted in later experiments).
-   One possible explanation is an effect of dose. While Chen et al., used a concentration of 1 g/l, we, unfortunately, tested 0.5 g/l, based on experience from previous studies, making a direct comparison difficult.
-   We did find significant over-expression of cytokines relevant in inflammation, indicating some effect of the treatments.
-   Since these biochemical signals were not strong enough to manifest phenotypically, we reasoned that our treatment wasn't sufficiently injurious.
-   We therefore decided to increase the dosage to 1 g/l for subsequent experiments, as well as to increase the osmolality of the formula for the next feeding experiment (Exp 3).
-   With assurance in our experimental setup we decided to include further measures of intestinal dysfunction.
-   We also decided to measure the birth weight of all pups, in addition to body weight at the beginning of the experiment (day 3) and end (day 5).

### Experiment 2: Effects of various types of antibiotics - (Microbiome dysbiosis)

#### Rationale:

-   Exp 1 demonstrated that the type of feed (breastmilk or formula) had a greater effect on cytokine expression than maternal antibiotics treatment.
-   Central to our hypothesis of NEC development is an intestinal microbiome dysibiosis, and a broad-spectrered antibiotic treatment potentially resulted in an overall depletion rather than the manifisation of a dysbiosis.
-   We hypothesized that targeting specific strains of bacteria in the mother would lead to a more dysbiotic microbiome and allow for greater subsequent dysbiosis in the offspring.

#### Aim:

-   To test the effect of gram-positive-specific antibiotics versus gram-negative-specific antibiotics, without formula feeding on intestinal permeability and cytokine expression.
-   To see whether a larger dose of antibiotics (1 g/l) causes decreased body weight at birth and at day 5 compared with non-treated controls.

#### Conclusion:

-   Birth weight: `r mean(bodyweight_birth_con` g (CON) vs `r mean(bodyweight_birth_mat` g (MAT)
-   Visual inspection: no clear signs of intestinal inflammation
-   Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 2.

*Antimals and experimental setup* - The breeding setup in Exp 2 was identical to Exp 1.
- On gestational day 15 vancomycin, gentamicin or control (tab water) was added to the drinking water pregnant females, forming 3 experimental groups; VANCO, GENTA and CON.

*Weight development* - We checked the cages twice daily from gestational day 18 and measured the weight of newly-born pups.
- VANCO and GENTA mothers gave birth on the 25th of february (getational day 20) with mean offspring body weights of `r mean(bodyweight_birth_vanco` (N = `r nrow(bodyweight_birth_vanco`) and `r mean(bodyweight_birth_genta` (N = `r nrow(bodyweight_birth_vanco`) respectively.
- The CON mother gave birth on the 26th of february (gestational day 21) with mean offspring body weight of `r mean(bodyweight_birth_con` (N = `r nrow(bodyweight_birth_vanco`).
- This difference in body weight at birth was statistically significant (p = `r lm_weight_exp2$p.value`), but adjusting for length of gestation erased the statistical significance (p = `r lm_weight_exp2_adj$p.value`).
- On day 3 and 5 this difference was in body weight was no longer detectable.
*(Figure of body weights in exp 2)*

*Intestinal permeability - FITC-Dextran administration* - On the day of euthanasia pups were separated from the mothers at 07:00 (halting breastfeeding).
- At 10:00, 50 µL FITC-dextran (0.045 mg/µL),for a final dose of 750 mg/kg, was administered orally using a PICC line and a 1 mL syringe.
- Two hours later euthanasia and sample collection began (12:00).
- Pups were euthanized by cervical dislocation and blood samples collected and spun in Eppendorf tubes.
- FITC molecules were detectable in all groups.
- The highest concentration was found in the GENTA group (`r fitc_exp2_genta` µg/ml) and the lowest in the VANCO group (`r fitc_exp2_vanco` µg/ml), respectively, with the CON group in between (`r fitc_exp2_vanco` µg/ml).
- This difference was statistically significant (p = `r anova_fitc_exp2`, one-way analysis of variance).
*(Figure of FITC in exp 2)*

*Cytokine expression* - Tissue weight-adjusted results of cytokine concentrations showed significant differences between the 3 groups for `r signif_cyto_exp2`.
### Rerun these experiments.

#### Commentary and interpretation experimental results - Exp 2.

-   Weight development was now similar to the findings of Chen et al.

-   Whether it was a result of growth restriction or shorter gestational age is difficult to determine.

-   Chen et al. did not report whether dams were from the same day.

-   Intestinal permeability response was different depending of type of antibiotics.

-   Vancomycin decreased intestinal permeability whereas gentamicin increased it.

    -   This has been shown in Winstar rats before (Thulstrup 2015).

-   Cytokine expression was most affected in the VANCO group.

-   In summary:

-   In experiment 2, Maternal antibiotics treatment did not have an effect.

-   This indicates that the greater effect is of formula-feeding, similar to the findings in exp 1.

-   The intestinal permeability assay is interesting in that it supports the findings of Thulstrup et al. 2015.

### Experiment 3 - Interaction effects between antibiotics treatment and formula feeding

#### Rationale:

Based on the results from Exp 2 it was difficult to ascertain which treatment was more likely to manifest in NEC like phenotype.
- Gentamicin had increased intestinal permeability the most.
- However, the greatest effect of cytokine expression was found with the vancomycin treatment.
- In addition to this divergence, the interaction effect found in Exp 1 suggested it was likely that combining MAT with formula feeding would yield different results.
- Therefore, we decided to increase the breeding population to allow for an experiment that combined both types of antibiotics with either formula-feeding or breastfeeding, for a total of 6 experimental groups.

#### Aim:

-   Repeat Exp 2 with greater sample sizes, adding the effect of formula feeding.
-   To test the effect of gram-positive-specific antibiotics versus gram-negative-specific antibiotics in combination with formula feeding on intestinal permeability and cytokine expression.
-

#### Conclusion:

-   Combining MAT with formula feeding resulted in more pronounced signs of inflammation than MAT alone.
-   VANCO-FF and VANCO-BF had the highest and lowest intestinal permeability scores, respectively.
-   Significantly altered expression of XXSignificantSCYTOKINESXX.
-   Greatest determinant of variation was formula-feeding.

#### Detailed description of Exp 3.

*Antimals and experimental setup* - To increase the number of pups available we doubled the number of mice for breeding, 36 females and 12 males.
- Pregnant females were again given either Vancomycin, Gentamicin (1 g/l) or sterile water on gestational day 15.
- `r nrow(preexp_3` pups were born and separated into breastfeeding or formula feeding, for a total of 6 groups: - VANCO-BF, VANCO-FF, GENTA-BF, GENTA-FF, CON-BF and CON-FF.

*Intestinal permeability - FITC-Dextran administration* - We repeated the protocol for FITC-administration, separating pups from mothers at 08:00 and dosing at 11:00.
- Formula-fed pups consistently showed higher levels of Serum FITC concentration compared to breastfed pups, irrespective of MAT status.
- VANCO-FF and VANCO-BF had the highest and lowest levels of FITC, respectively.

*Cytokine expression* - We found significant differences in concentrations of XXSignificantSCYTOKINESXX.
Linear regression analysis indicated the biggest determinant of differences in outcomes was formula feeding, with the type of antibiotic being of lesser importance compared.
- Directionally the trend was the same for all cytokines.
- IL-1b: Non-significant decrease in all antibiotics-treated offspring, irrespective of feeding.
- IL-2: significantly lower concentrations in all formula-fed pups compared with CON-BF.
- IL-6: Significantly greater levels in all formula-fed pups compared with CON-BF.
- INFg: significantly lower concentrations in all formula-fed pups compared with CON-BF.
- KCGRO: Significantly greater levels in all formula-fed pups compared with CON-BF.

#### Commentary and interpretation experimental results Exp 3:

-   Intestinal permeability was by far the greatest in VANCO-FF, with levels of FITC-dex close to 10-fold increased in this group.
-   In contrast, VANCO-BF was the group with the lowest measure of intestinal permeability, similar to the finding in exp 2.
-   Looking at **FIGURE CYTOKINES ROUND 3** shows how formula-fed groups systematically over- or underexpress compared to breastfed analimals, irrespective of type of antibiotic, indicating that the biggest determinant of cytokine variation was again formula feeding.

*Adjusting for weight of tissue* - In the previous experiments, cytokine concentrations were adjusted for the weight of the tissue analyzed, to control for biopsy sizes.
- We did so because the average weight of tissue samples was not meaningfully different between groups.
- This assumption wasn't met in Exp 3, where samples from formula-fed pups weighted on average XX%XX more (p = XXp.valueXX) than breastfed controls.
- Vanco-FF had the heaviest ileum tissue samples of all **mean + P (if its true)**.
- Tissue edema could well be indicative of inflammation, and hence desirable, however it introduces a bias against the formula-fed pups if cytokine concentrations were normalized to weight.
- The reported findings are no longer significant **Table of significant cytokines** when adjusting for weight.

-   This phenomenon is a common challenge in animal disease models, requiring measure of normalization other that tissue weight, such as total protein content.
-   Unfortunately this discovery was done after tissue homogenization and analysis, making it too late for further analysis.

*Combining results from exp 2 and exp 3* - The protocols for the breastfed groups were identical across experiment 2 and 3, allowing for pooling of these results.
- While analyzing the results from both groups together increased the group sizes and degrees of statistical freedom, it did not change the clinical conclusions.

### Experiment 4 - FVT Trial

#### Rationale: **What was the actual rationale?**

-   Central to the development of NEC is an assumption of a gut microbial dysbiosis, which we tried to mimick by administering antibiotics prenatally.
-   Results from exp 1-3 indicated that maternal antibiotics treatment had little effect in the context of breastfeeding, but when combined with formula-feeding was able to induce mild symptoms of intestinal inflammation.
-   It was unclear whether this was solely an effect of the formula, or whether an interaction with the maternal antibiotics treatment existed.
-   If our findings were partly due to a successful manipulation of the gut microbiome, then eliminating the induced disbyosis should also partly mitigate the inflammatory signatures.
-   Relative bacterial abundance / bacterial composition is partly a function of viral (bacteriophage) abundance, and sevaral studies have shown how a FVT can alter the bacterial composition.
-   We therefore decided to conduct a final experiment, in which we administered an FVT to a vancomycing-treated, formula-fed group (since this combination (VANCO-FF) had the biggest effects in terms of intestinal permeability and proinflammatory cytokine expression in Exp 3).
-   We hypothesized that any dysbiosis created by maternal antibiotics treatment could be mitigated with a follow-up administration of FVT prior to beginnig formula-feeding.
-   If successful, the FVT-treated group would show fever intestinal injury scores compared with a non-FVT treated group.
-   Previous experiments had shown that MAT in combination with breastfeeding had no effect on inflammatory markers and VANCO-BF thus acted as a healthy control to both formula-fed treatment groups. **(The problem here is that formula is the only thing with an effect. A breastfed control only shows the effect of the formula)**

#### Aim:

-   To test whether signs of inflammation are reduced in pups treated with an FVT.
-   To test whether any reductions in symptoms can be explained with a manipulation of the microbial composition.

#### Conclusion:

-   FVT treated pups showed similar intestinal permeability and pro-inflammatory cytokine concentrations as non-FVT-treated controls.
-   An `r fold_increase_reg3g`-fold increase in expression of the `r REG3G` gene was found in the FVT group.

#### Detailed description of Exp 4.

-   For the final experiment we repeated the previous breeding setup, assigning 36 females and 12 males for breeding.

-   Only 5 of the 36 females became pregnant, and all were assigned to vancomycin (1 g/l) from pregnancy-day 15 until delivery, giving birth to `r nrow(preexp_4` pups on Saturday the XXth.

-   2 days after birth pups were separated into breastfeeding or formula feeding.

-   The formula-fed pups were further divided into a treatment and control group, receiving either 30 µl fecal virome filtrate (FVT) (0.01 mg farces/µL) or an equal amount of SM buffer (control) 1 hour prior to the first feeding.

-   We repeated the FVT administration 24 hours later (09:00 day 2) for a total dose of 0.6 mg feces material per pup.

-   2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration.
    Intestinal samples were collected for gene expression assays, histological assessment, pro-inflammatory cytokine analysis and gut microbiome sequencing.

*Intestinal permeability - FITC-Dextran administration* - We repeated the protocol for FITC-administration, separating pups from mothers at 08:00 and dosing at 11:00.
- The FVT and FVT-control (SM) had similar permeability scores.
- Compared with the vancomycin-treated formula-fed pups from experiment 3, there were no significant differences **Figure X**.
- The breastfed pups did, however, have significantly higher levels compared with all other breastfed pups.

*Cytokine expression* - We found significant differences in concentrations of TNFa in Exp 4, with greater level in the FVT-treated group (XXX), compared to both VANCO-BF and VANCO-FF (SM) (XXX and XXXX respectively).
- This is in contrast experiment 3, where differences between VANCO-BF and VANCO-FF found significant differences for XXXCYtokinesXXX.
OR: - This is similar to the findings of experiment 3, where no differences between VANCO-BF and VANCO-FF were found either.

*Gene expression* - qPCR analyzis of intestinal samples found statistically significant differences in the level of XXXSIGNIFICANT_GENESXXX expression between groups with the delta-delta method (reformulate).
- However, a difference in expression of the reference gene between the FVT and breastfed control was observed.
this difference was statistically significant, invalidating the assumption of *equal baseline expression.*(reformulate) - Comparing the groups on the basis of (SOMEHTING ELSE THAN DELTA DELTA) found a `r fold_increase_reg3g`-fold increased expression of the `r REG3G` gene in the FVT groups.

#### Commentary and interpretation

-   FVT-treatmed did not seem to change the intestinal permeability.

-   It is possible that something went wrong with the VANCO-BF group, given FITC-concentration was `r mean(fitc_level_exp4$bf`, which is 4-5 times higher than VANCO-BF groups on identical protocols in Exp 2 `r mean(fitc_level_exp2$bf` and 3 `r mean(fitc_level_exp3$bf`. Compared with theses levels the development looks more biologically reasonable. **FIGURE of EXP4 with DIFFERENT CONTROL GROUP**

-   The level of pro-inflammatory cytokines were the same in the FVT and SM groups, again indicating a lack of efficacy of the FVT treatment.

-   Summary: The FVT treatment did not seem to have an effect.

### \--\>Post experimental analyses and Discussion of results

-   Histological image analysis was performed after all experiments had finished.

-   We inspected the morphology of the ileum and colon, and visually assessed thickness of the mucin layers.

-   Analyzing AB-PAS stains for mucin-stained tissue to the total area, showed no differences between any treatment groups.

-   Formula-fed and breastfed pups were slightly different, irrespective of antibiotics treatment, whereas there were no observable differences between antibiotic treatments.

-   These findings supports the visual inspections of intestines made during sampling, where no signs of inflammation were detected.

-   *Images of intestines*

-   Gut microbiome composition

## Discussion