### 2022-11-04 submission/thesis.Rmd

---
title: " "
output:
  pdf_document:
    number_sections: TRUE
geometry: "left = 2.5cm, right = 2cm, top = 2cm, bottom = 2cm"
fontsize: 12pt
mainfont: Arial 
header-includes:
  - \usepackage{float}
  - \usepackage{sectsty}
  - \usepackage{paralist}
  - \usepackage{setspace}\spacing{1.5}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
  - \usepackage{dcolumn}
  - \usepackage{natbib}\bibliographystyle{agsm}
  - \usepackage[nottoc, numbib]{tocbibind}
  #- \usepackage[english, sund, titlepage, hyperref]{ku-frontpage} (Consider using  #extra_dependencies)
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

\allsectionsfont{\centering}
\subsectionfont{\raggedright}
\subsubsectionfont{\raggedright}

\pagenumbering{gobble}

```{r uni_logo, echo=F, out.width="95%"}
knitr::include_graphics("img/sund.pdf")
```

\newpage

\begin{centering}

\vspace{1cm}

\Large
{\bf The University Of Edinburgh}

\Large
{\bf School Of Geosciences}

\vspace{1cm}

\Large

\doublespacing
{\bf COMPARISON OF TOP-DOWN AND BOTTOM-UP APPROACHES ON SPECIFIC LEAF AREA PATTERNS, \\AT GLOBAL, LATITUDINAL, AND BIOME SCALES}

\vspace{1 cm}

\normalsize
\singlespacing
By

\vspace{0.5 cm}

\Large

{\bf ANNA CHIRUMBOLO}

\vspace{1.5 cm}

in partial fulfilment of the requirement \\for the degree of BSc with Honours \\in Ecological and Environmental Sciences

\vspace{1.5 cm}

\normalsize
mm yy

\end{centering}

\newpage

\pagenumbering{gobble}

\begin{centering}

{\bf Abstract}

\end{centering}

\spacing{1.5}

(the spacing is set to 1.5)

no more than 250 words for the abstract

- a description of the research question/knowledge gap – what we know and what we don’t know
- how your research has attempted to fill this gap
- a brief description of the methods
- brief results
- key conclusions that put the research into a larger context

\pagenumbering{roman}

\newpage

\centering
\raggedright
\newpage
\tableofcontents

\newpage

```{r acknowledgments, child='acknowledgments.Rmd'}
```

\newpage

```{r intro, child = 'introduction.Rmd'}
```
\pagenumbering{arabic}

\newpage

```{r methods, child = 'methods.Rmd'}
```

\newpage

```{r results, child= 'results.Rmd'}
```

\newpage

```{r discussion, child = 'discussion.Rmd'}
```

\newpage

```{r conclusion, child = 'conclusion.Rmd'}
```

\newpage

```{r biblio, child = 'bibliography.Rmd'}
```

\newpage

```{r appendix, child = 'appendix.Rmd'}
```