### 2022-11-08 submission/appendix.Rmd

---
title: ''
output: pdf_document
---
# Appendix(ces)

## Appendix A: additional tables

\newpage

## Appendix B: additional figures

\newpage

## Appendix C: code

```{r get-labels, echo = FALSE}
labs = knitr::all_labels() # this is the new code chunk to add
labs = setdiff(labs, c("setup", "get-labels")) # this function excludes the code chunk with the label "setup" from being displayed.
```

```{r all-code, ref.label=labs, eval=FALSE}
# this code chunk displays all source code from your entire dissertation document (that you have written in .Rmd, not from your R script).
```