### 2022-11-11 code/weight_analysis.Rmd

---
title: "Weight analysis"
author: "Kaare D. Tranaes"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstatix)
library(ggpubr)
library(patchwork)
library(tidyverse)
library(readxl)
library(broom)
```

#___Analysis of the weight development in the MARTRINEM study___
```{r}
data <- read_excel("data/processed/weight_development.xlsx") %>% 
  select(sample_id:cage, bodyweight_birthday_group_average, bodyweight_birthday_plus_24H_group_average, bodyweight_birthday_plus_48H_group_average, bodyweight_baseline, bodyweight_baseline_plus_24h, bodyweight_baseline_plus_48h) %>% 
  rename(day_0 = "bodyweight_birthday_group_average",
         day_1 = "bodyweight_birthday_plus_24H_group_average", 
         day_2 = "bodyweight_birthday_plus_48H_group_average",
         day_3 = bodyweight_baseline, 
         day_4 = `bodyweight_baseline_plus_24h`, 
         day_5 = `bodyweight_baseline_plus_48h`,) %>% 
  pivot_longer(., cols = c(day_0:day_5),
                names_to = "time",
                values_to = "bodyweight") %>% 
  mutate(received_antibiotics = recode_factor(received_antibiotics, 
                                       Yes = "MAT",
                                       No = "CON"))
  #%>%
  # #mutate(across(where(is_character),as_factor))
```

#_bodyweight in all experimets._
```{r}
startweight <- data %>% 
  filter(exp_number %in% c("1", "2", "3","4")) %>% 
  filter(time %in% c("day_3", "day_5")) %>%
  drop_na(bodyweight) %>% 
  group_by(exp_number, type_of_feed, maternal_treatment, group, time) %>% 
  summarise(startweight = mean(bodyweight),
            sd = sd(bodyweight))

startweight %>% 
  filter(exp_number == "2") %>% 
  arrange(time)
```

#_birthweight in exp 2_
```{r}
birthweight <- data %>% 
  filter(exp_number %in% "2") %>% 
  filter(time %in% c("day_0", "day_1", "day_2")) %>%
  drop_na(bodyweight) %>% 
  group_by(exp_number, maternal_treatment, time) %>% 
  summarise(birthweight = mean(bodyweight),
            sd = sd(bodyweight))

lm_birthweight <- data %>%  #LM for effect of gestational age on offspring birthweight. 
  filter(exp_number %in% "2") %>% 
  filter(time %in% c("day_0", "day_1", "day_2")) %>%
  drop_na(bodyweight) %>% 
  filter(time == "day_0") %>% 
  mutate(gestation_age = if_else(maternal_treatment == "CON", 21,20)) %>% 
  do(lm(bodyweight ~ gestation_age,.) %>% tidy)
lm_birthweight

data %>% # ANOVA for effect of gestational age on offspring birthweight. 
  filter(exp_number %in% "2") %>% 
  filter(time %in% c("day_0", "day_1", "day_2")) %>%
  drop_na(bodyweight) %>% 
  filter(time == "day_0") %>% 
  mutate(gestation_age = if_else(maternal_treatment == "CON", 21,20)) %>% 
  rstatix::anova_test(., bodyweight ~ gestation_age)

```

#_Birthweight of offspring, experiment 2_
```{r}
exp2.bw.density <- data %>% 
  filter(exp_number == "2") %>% 
  filter(time == 'day_0') %>% 
  ggplot(aes(x = bodyweight, color = received_antibiotics)) +
    geom_density() +
  geom_vline(data = data %>% 
              filter(exp_number == "2") %>% 
  filter(time == 'day_0') %>% 
  group_by(received_antibiotics) %>% 
  summarise(grp.mean = mean(bodyweight)), aes(xintercept=grp.mean, color=received_antibiotics),
             linetype="dashed") +
  guides(fill = guide_legend(title = ""),
         colour = guide_legend(title = " ")) + 
  labs(title = "", y = "n", x = "Bodyweight (g)") +
  theme_minimal()

exp2.bw.col <- data %>% 
  filter(exp_number == "2") %>% 
  filter(time == 'day_0') %>% 
  ggplot(aes(x = received_antibiotics, y = bodyweight, fill = received_antibiotics)) +
  geom_bar(stat = "summary",
           fun = "mean") +
  labs(title = "", y = "Mean body weight", x = "") +
  theme_minimal() +  
  stat_compare_means(
    method = "t.test", 
    label.y = 1.6, 
    label.x.npc = 0.4) +
  theme(legend.position="none")

birth_weight_patch <- (exp2.bw.density + exp2.bw.col) + plot_annotation(
  title = 'Effect of Maternal Antibiotics Treatment on birth weight of offspring',
  subtitle = 'Experiment 2',
  tag_levels = 'A',
  caption = 'MAT = Maternal Antibiotics Treatment, CON = Water Controls \n 
  Statistics: non-paired two.sided t-test.'
)
birth_weight_patch
#ggsave("../Results/Weight/Figure_1_Birthweights.png", birth_weight_patch)

```

#_Weight development during experiments_
```{r eval=FALSE, include=FALSE}
bw_development_feed <- data %>% 
filter(exp_number %in% c("1","3","4")) %>% 
filter(time %in% c("day_3", "day_5")) %>%
ggplot(aes(x = time ,
           y = bodyweight, 
           fill = type_of_feed)) +
  facet_wrap(~ exp_number + type_of_feed, 
             ncol = 2) + 
  geom_bar(stat = "identity",
           position = "dodge") + 
  guides(fill = guide_legend(title = " ")) +
  labs(
      caption = "Experimental round 2 not shown due to no formula feeding.",
    y = "Bodyweight (g)"
  ) +
  theme_bw()

bw_development_ab <- data %>% 
filter(exp_number %in% c("1","2","3","4")) %>% 
filter(time %in% c("day_3", "day_5")) %>%
ggplot(aes(x = time ,
           y = bodyweight, 
           fill = received_antibiotics)) +
  facet_wrap(~ exp_number + received_antibiotics, 
             ncol = 2) + 
  geom_bar(stat = "identity",
           position = "dodge") + 
  guides(fill = guide_legend(title = " ")) +
  labs(caption = "All mice recieved antibiotics in experimental round 4.",
    y = "Bodyweight (g)"
  ) +
  theme_bw()

bodyweight_develop_patch <- (bw_development_feed + bw_development_ab) + plot_annotation (
  title = 'Effect of Maternal Antibiotics Treatment on body weight development of offspring',
  subtitle = 'Faceted by experiment and Feeding or MAT-status',
  tag_levels = 'A',
  caption = 'Form = formular feed, Milk = breastfeeding.  
      MAT = Maternal Antibiotics Treatment, CON = Water Controls.')

bodyweight_develop_patch
#ggsave("../Results/Weight/Figure_2_bodyweight_development.png", bodyweight_develop_patch)

```

### Research Question 1: Body weight of offspring from water and antibiotic-treated mice at baseline

```{r style baseline bodyweight}
get_box_stats <- function(y, upper_limit = max(data$Bodyweight[data$Time == "Day2"]) * 1.0) {
  return(data.frame(
    y = 0.9 * upper_limit,
    label = paste(
      "n =", length(y), "\n",
      "Mean =", round(mean(y), 2), "\n" #,
      #"Median =", round(median(y), 2), "\n"
    )
  ))
}

my_comparisons <- list( c("Water", "AB"))

data %>% 
  filter(Time == "Day2") %>% 
  ggplot(aes(
    received_antibiotic, Bodyweight
  )) + 
  geom_boxplot() +
  facet_wrap(~experiment_no, 
             nrow = 1) + 
  geom_dotplot(
     binaxis = "y", stackdir = "center", dotsize = 0.5) + 
  stat_summary(
    fun.data = get_box_stats, geom = "text", hjust = 0.5, vjust = 0.9) +
  labs(
    title = "Body weight of offspring from antibiotic-treated vs water-treated mice",
    subtitle = "Results shown from each of the 4 experimental rounds",
    caption = "In Exp. 4 all mice received antibiotoics, hence no control.
    Comparison method: Student T-test",
    x = "Maternal Treatment",
    y = "Bodyweight at end of study (g)"
  ) +
  theme_bw() + 
  stat_compare_means(
    method = "t.test", 
    label.y = 3, 
    comparisons = my_comparisons
    )
```

```{r weightgain plots}
############# Effect of Formula vs. Breastfeeding on Weight Gain" ###################

bodyweight_gain_BFvsFORM <- data %>% 
  mutate(weightgain = weight_end - weight_baseline) %>% 
  ggplot(aes(
    type_of_feed, weightgain
  )) + 
  geom_boxplot() +
  facet_wrap(~experiment_no, 
             nrow = 1) +
geom_dotplot(
     binaxis = "y", stackdir = "center", dotsize = 0.5) + 
  stat_summary(
    fun.data = get_box_stats, geom = "text", hjust = 0.2, vjust = 0.9) +
  labs(
    title = "Effect of Formula vs. Breastfeeding on Weight Gain",
    subtitle = "Results shown from each of the 4 experimental rounds",
    caption = "In Exp. 2 all mice were breastfed.
    Comparison method: Student T-test",
    x = "Type of feed",
    y = "Body weight increase  (g)"
  ) +
  theme_bw() + 
  stat_compare_means(
    method = "t.test", 
    label.y = 3, 
    comparisons = my_comparisons_2
    )
my_comparisons_2 <- list( c("Breastmilk", "Formula"))
bodyweight_gain_BFvsFORM
```

```{r weightgain plot 2}
############# Effect of  Antibiotocis on weightgain in Breastfed offspring" ################

bodyweight_gain_WATER_BF_vs_AB_BF <- 
  data %>% 
  mutate(weightgain = weight_end - weight_baseline) %>% 
  filter(type_of_feed == "Breastmilk") %>% 
  ggplot(aes(
    received_antibiotic, weightgain#, colour = mother
    )) + 
  geom_boxplot()+
  geom_point()+
  facet_wrap(~experiment_no, 
             nrow = 1) +
  labs(
    title = "Breastfed offspring +/- Maternal Antiobiotics Treatment"
  )
bodyweight_gain_WATER_BF_vs_AB_BF
```