### 2022-11-12 code/Round 3/FITC round 2 and 3 copy.Rmd

---
title: "FITC analysis script - round 2 and 3"
author: "Kaare"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

```{r load packages, include=FALSE}
library(readxl) 
library(tidyverse)
library(gt)
library(gtsummary)
library(broom)
library(rstatix)
library(knitr)
library(patchwork)
library(ggpubr)
```

# Introduction to script

This script is useful for analyzing intestinal permeability from FITC-Dextran assays.

Required data structure is a wide tibble (column headers = variables, samples and values = rows).

This script is per default set to handle ileum and serum expression, although any tissue will work. This will just require resetting of the "tissue"-variable.

# FITC for experiment 3.

### Basic data structure and results table.

First you load the data and check the structure. Make sure cytokine concentrations are all numeric (<dbl>). If not, mutate non-numeric columns with mutate_if(is_character).

```{r load data, echo=FALSE}
# Import dataset
fitc <- read_excel("data/processed/fitc.xlsx") %>% 
  rename(values = `result_µg/ml_undiluted`) %>%  #rename undiluted column.
  drop_na(values)

glimpse(fitc)
```

Set the experimental round here:

```{r load data and set tissue type, include=FALSE}
#set tissue type
study_round <- c("2","3")
#study_round <- 3
```

Check for data distribution.

```{r load data and set tissue type, include=FALSE}
#set tissue type

fitc %>% 
  filter(exp_number == study_round) %>% 
  group_by(exp_number, group2) %>% 
  shapiro_test(values)
#  group2   variable statistic           p
#  <chr>    <chr>        <dbl>       <dbl>
# CON-BF   values       0.975 0.946      
# CON-FF   values       0.912 0.295      
# GENTA-BF values       0.965 0.838      
# GENTA-FF values       0.962 0.833      
# VANCO-BF values       0.854 0.0155     
# VANCO-FF values       0.526 0.000000232

fitc %>%  # try log-tranformation
  filter(exp_number == study_round) %>% 
  group_by(group2) %>% 
  mutate(values = log10(values)) %>% 
  shapiro_test(values)

#group2   variable statistic         p
#  <chr>    <chr>        <dbl>     <dbl>
# CON-BF   values       0.877 0.0640   
# CON-FF   values       0.955 0.731    
# GENTA-BF values       0.851 0.0445   
# GENTA-FF values       0.943 0.636    
# VANCO-BF values       0.833 0.00778  
# VANCO-FF values       0.695 0.0000171   # Log transforming doesn't help.

#Density plots. Works
fitc %>% 
  filter(exp_number == study_round) %>% 
  ggplot(aes(x = values, fill = group2)) + 
  geom_density() +
  facet_wrap(~ group2,
             scales = "free")

#Jitter plots. Works
fitc_long %>% 
  filter(exp_number == study_round) %>% 
  ggplot(aes(x = group2, y = values)) + 
  geom_jitter() +
  facet_wrap(~ group2,
             scales = "free")

```

Exclude outliers
```{r data distribution}
outliers <- fitc %>% 
  filter(exp_number == study_round) %>% 
  group_by(group2) %>% 
  mutate(values = ifelse(values > 50, mean(values), values)) %>% 
qplot(x = group2, y = values, data = ., geom = "jitter",
    color = group2)

```

Below is a table of the levels of intestinal permeability measurable as the concentration of fitc-dextran in serum for each of the experimental groups.

The values will be displayed as mean + sd if normally distributed or median + IQR if not. An appropriate statistical test for differences will be performed (student-t for two, normally distributed groups, Anova for more than 2 groups, Kruskal wallis for non-normally distrubuted results.)

```{r}
tbl1 <- fitc %>% 
  filter(exp_number == 2 | exp_number == 3) %>% 
  mutate(Log10_values = log10(values)) %>% 
  select(group2, values, Log10_values) %>% 
  tbl_summary(by = group2,
              digits = everything() ~ c(2,1,1),
              missing = "no",
              statistic = all_continuous() ~ "{mean} ({sd})") %>%
  add_p(values ~ "aov") %>% 
  modify_header(
    label = "FITC Conc.",
    stat_1 = "**CON-BF**  \n (n = {n})",
    stat_2 = "**CON-FF**  \n (n = {n})",
    stat_3 = "**GENTA-BF**  \n (n = {n})",
    stat_4 = "**GENTA-FF**  \n (n = {n})",
    stat_5 = "**VANCO-BF**  \n (n = {n})",
    stat_6 = "**VANCO-FF**  \n (n = {n})") %>% 
  modify_spanning_header((c("stat_1", "stat_2", "stat_3","stat_4", "stat_5", "stat_6") ~ "Treatment groups (N total = {N})"))

tbl1_gt <- tbl1 %>% 
 as_gt() %>%
  tab_header(title = md("**Intestinal permeability**")) %>%
  opt_align_table_header("center") %>% 
  tab_options(heading.title.font.size = 20,
              heading.background.color = "lightgrey",
              ) %>% 
  tab_source_note(
    source_note = " * P-value indicates overall group differences"
  ) %>%
  tab_source_note(
    source_note = md("**Matrinem round 3**")
  ) %>%
  cols_width(label ~ px(100),
             stat_1 ~ px(150),
             stat_2 ~ px(150),
             stat_3 ~ px(150),
             stat_4 ~ px(150),
             stat_5 ~ px(150),
             stat_6 ~ px(150))

tbl1_gt
# Save and export files
# tbl1 %>%  #Save as excel file.
#   as_hux_table() %>%
#   huxtable::quick_xlsx(., file = "../Results/Round 1/ileum-cytokines.xlsx")
# 
# tbl1 %>% #Save as image file.
# as_flex_table() %>%
#   flextable::save_as_image(path = "../../Results/Round 2/ileum_cytokines_table.png",webshot = "webshot")
```

This table can be exported with the following code either to word, excel, pdf, html or image files. See the syntax of huxtable and flextable (?flextable).

```{r ileum cytokine export, echo=TRUE, warning=FALSE}
tbl1_p <- tbl1 %>%
  filter_p() %>% 
  as_tibble()
```

### Plots and regression models

First the data must be transformed to a long format (pivot longer). Mutate any new columns with group treatments needed for the regression model (in this case MAT and FEED (experimental factors)).

```{r long format, echo=FALSE}
fitc_long <- fitc %>% 
  drop_na(
    values) %>% 
  mutate(
    group = recode_factor(  #Change the order of the groups to desired plot order
      group,
      "CON-BF"= "CON-BF",
      "CON-FF" = "CON-FF",
      "GENTA-BF"= "GENTA-BF",
      "GENTA-FF"= "GENTA-FF",
      "VANCO-BF"= "VANCO-BF",
      "VANCO-FF"= "VANCO-FF"
      ))

head(fitc_long)

```

Create a comparison vector to see pariwise comparisons

```{r plots all fitc, warning=FALSE}
my_comparisons <- list( c("CON-BF", "CON-FF"), c("GENTA-BF", "GENTA-FF"), c("VANCO-BF", "VANCO-FF"), c("GENTA-FF", "VANCO-FF"))

plot_all_fitc <- fitc_long %>%    #____Normal values_____#
  filter(exp_number == study_round) %>% 
ggplot(aes(x = group2, 
           y = values, 
           fill = group2)) + 
  geom_boxplot() +
  geom_point() +
  #coord_cartesian(ylim = c(1.5, 16)) + 
  stat_compare_means(comparisons = my_comparisons, label.y = c(30, 30, 30, 50), method = "wilcox.test")+
  # stat_compare_means(aes(label = paste0("Kruskal-Wallis \n p = ", ..p.format..)),
  #   label.y.npc = 0.75,
  #   label.x.npc = 0.55) +
  labs( # Change title, and axis names
       x = "",
       y = "Serum FITC conc. (µg/ml)",
       fill = "Groups",
       caption = "Matrinem round 3") +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="lightgrey"),#Change legend box background color
    plot.caption = element_text(color = "darkgrey", size=5)) + 
  scale_fill_brewer(palette="Set1")

plot_all_fitc_log <- fitc_long %>%        #Log_transformed values_____#
  filter(exp_number == study_round) %>% 
ggplot(aes(x = group2, 
           y = log10(values), 
           fill = group2)) + 
  geom_boxplot() +
  geom_point() +
  #coord_cartesian(ylim = c(1.5, 16)) + 
  stat_compare_means(comparisons = my_comparisons, label.y = c(1.5, 1.5, 1.5, 1.7), method = "wilcox.test")+
  # stat_compare_means(aes(label = paste0("Kruskal-Wallis \n p = ", ..p.format..)),
  #   label.y.npc = 0.75,
  #   label.x.npc = 0.55) +
  labs( # Change title, and axis names
       x = "",
       y = "Log10 serum FITC conc. (µg/ml)",
       fill = "Groups",
       caption = "Matrinem round 3") +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="lightgrey"),#Change legend box background color
    plot.caption = element_text(color = "darkgrey", size=5)) + 
  scale_fill_brewer(palette="Set1")

all_fitc_patch <- plot_all_fitc + plot_all_fitc_log + plot_annotation (
  title = 'Intestinal permeability',
   #subtitle = 'All fitc measured',
  tag_levels = 'A',
  caption = 'CON = Water, BF = Breast-Feeding, GENTA = Gentamicin,  VANCO = Vancomycin,  \n
  Pariwise comparisons = Wilcoxon test.',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"),
                plot.subtitle = element_text(hjust = 0.5, face = "bold.italic"),
                plot.caption = element_text(hjust = 0.5, size=7,face = "bold")))

all_fitc_patch
#ggsave('../../Results/Round 1/All fitc-boxplot.png',all_fitc_patch)
```