### 2022-11-13 code/Round 4/FITC round 4.Rmd

---
title: "FITC analysis script - round 4"
author: "Kaare"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

```{r load packages, include=FALSE}
library(readxl) 
library(tidyverse)
library(gt)
library(gtsummary)
library(broom)
library(rstatix)
library(knitr)
library(patchwork)
library(ggpubr)
```

# Introduction to script

This script is useful for analyzing intestinal permeability from FITC-Dextran assays.

Required data structure is a wide tibble (column headers = variables, samples and values = rows).

This script is per default set to handle ileum and serum expression, although any tissue will work. This will just require resetting of the "tissue"-variable.

# FITC for experiment 4.

### Basic data structure and results table.

First you load the data and check the structure. Make sure cytokine concentrations are all numeric (<dbl>). If not, mutate non-numeric columns with mutate_if(is_character).

```{r load data, echo=FALSE}
# Import dataset
fitc <- read_excel("data/processed/fitc.xlsx") %>% 
  rename(values = `result_µg/ml_undiluted`) %>% 
  drop_na(values)

glimpse(fitc)
```

Set the tissue type and experimental round here:

```{r load data and set tissue type, include=FALSE}
#set tissue type
study_round <- "4"
```

Check for data distribution.

```{r load data and set tissue type, eval=FALSE, include=FALSE}
#set tissue type

fitc %>% 
  filter(exp_number == study_round) %>% 
  group_by(group) %>% 
  shapiro_test(values) %>% 
  filter(p < 0.05)

fitc<- fitc %>% 
  mutate(values = log10(values))     # Log transforming doesn't help.

#Density plots. Works
fitc %>% 
  filter(exp_number == 4) %>% 
  ggplot(aes(x = values, fill = group)) + 
  geom_density() +
  facet_wrap(~ group,
             scales = "free")

#Jitter plots. Works
fitc %>% 
  filter(exp_number == 4) %>% 
  ggplot(aes(x = group, y = values)) + 
  geom_jitter() +
  facet_wrap(~ group,
             scales = "free") 
```

Exclude outliers
```{r data distribution, eval=FALSE, include=FALSE}
outliers <- fitc %>% 
  filter(exp_number == study_round) %>% 
  group_by(group2) %>% 
  mutate(values = ifelse(values > 50, mean(values), values)) %>% 
qplot(x = group2, y = values, data = ., geom = "jitter",
    color = group2)

```

Below is a table of the levels of intestinal permeability measurable as the concentration of fitc-dextran in serum for each of the experimental groups.

The values will be displayed as mean + sd if normally distributed or median + IQR if not. An appropriate statistical test for differences will be performed (student-t for two, normally distributed groups, Anova for more than 2 groups, Kruskal wallis for non-normally distrubuted results.)

```{r}
tbl1 <- fitc %>% 
  filter(exp_number == 4) %>% 
  #mutate(Log10_values = log10(values)) %>% 
  select(group, values) %>% 
  tbl_summary(by = group,
              digits = everything() ~ c(2,1,1),
              missing = "no",
              #type = list(values ~ 'continuous2', Log10_values ~ 'continuous2'),
              statistic = list(values ~ "{median} ({p25}, {p75})"),
              label = values ~ "µg/mL") %>%

  add_p() %>% 
  # modify_table_styling(
  #   columns = label,
  #   rows = label == "µg/mL",
  #   footnote = "Non-parametric distribution"
  # ) %>% 
  modify_header(
    label = "FITC Conc.",
    stat_1 = "**VANCO-BF** \n
    (n = {n})",
    stat_2 = "**VANCO-FVF** \n
    (n = {n})",
    stat_3 = "**VANCO-SM** \n
    (n = {n})") %>% 
  modify_spanning_header((c("stat_1", "stat_2", "stat_3") ~ "Treatment groups (N total = {N})")) %>%  
  bold_labels()

tbl1_gt <- tbl1 %>% 
 as_gt(include = -tab_footnote) %>%
  tab_footnote( 
    footnote = md("Median (IQR), Kruskal-Wallis."),
    locations = cells_column_labels(columns = c(p.value))
  ) %>% 
  tab_header(title = md("**Intestinal permeability**")) %>%
  opt_align_table_header("center") %>% 
  tab_options(heading.title.font.size = 20,
              heading.background.color = "lightgrey"
              ) %>% 
  # tab_source_note(
  #   source_note = " µg/mL values"
  # ) %>%
  tab_source_note(
    source_note = md("**Matrinem round 4**")
  ) %>%
  cols_width(label ~ px(100),
             stat_1 ~ px(150),
             stat_2 ~ px(150),
             stat_3 ~ px(150))
tbl1_gt
# tbl1_gt %>% 
#   gtsave("../../Results/Round 4/FITC table round 3 TEST.png", expand = 50) #use .tex for latex files
```

### Plots and regression models

First the data must be transformed to a long format (pivot longer). Mutate any new columns with group treatments needed for the regression model (in this case MAT and FEED (experimental factors)).

```{r long format, echo=FALSE}
fitc_long <- fitc %>% 
  filter(exp_number == 4)%>% 
  drop_na(
    values) %>% 
    mutate(Log10_values = log10(values)) %>% 
  mutate(
    group = recode_factor(#Change the order of the groups to desired plot order
      group,
      "CON-BF"= "VANCO-BF",
      "FVT-FORM" = "VANCO-FVF",
      "SM-FORM" = "VANCO-SM"
      ))

head(fitc_long)

```

Create a comparison vector to see pariwise comparisons.
```{r comparison list, warning=FALSE}
my_comparisons <- list(c("VANCO-FVF", "VANCO-SM"), c("VANCO-BF", "VANCO-FVF"))
```

Boxplots with jitter - normal values
```{r plots all fitc normal values, warning=FALSE}
plot_all_fitc <- fitc_long %>%
ggplot(aes(x = group, 
           y = values, 
           fill = group)) + 
  geom_boxplot() +
  geom_point() +
  coord_cartesian(ylim = c(10,35)) + 
  stat_compare_means(comparisons = my_comparisons, label.y = c(20, 25), method = "wilcox.test") +
  #stat_compare_means(aes(label = paste0("Kruskal-Wallis \n p = ", ..p.format..)),
   # label.y.npc = 0.75,
    #label.x.npc = 0.55) +
  labs( # Change title, and axis names
       x = "",
       y = "Serum FITC conc. (µg/ml)",
       fill = "Groups"
       #caption = "Matrinem round 4"
       ) +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="lightgrey"),#Change legend box background color
    plot.caption = element_text(color = "darkgrey", size=5)) + 
  scale_fill_brewer(palette="Set1")

all_fitc_patch <- plot_all_fitc + plot_annotation (
  title = 'Intestinal permeability',
   #subtitle = 'All fitc measured',
  #tag_levels = 'A',
  caption = 'FVF = Fecal Virome Filtrate, SM = SM-buffer  \n
  Pairwise comparisions using wilcoxon', 
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"),
                plot.subtitle = element_text(hjust = 0.5, face = "bold.italic"),
                plot.caption = element_text(hjust = 0.5, size=7,face = "bold")))

all_fitc_patch
#ggsave('../../Results/Round 3/FITC boxplots.png',all_fitc_patch)
```

Log transformed values
```{r plots all fitc log transformed, warning=FALSE}

plot_all_fitc_log10 <- fitc_long %>%
ggplot(aes(x = group, 
           y = Log10_values, 
           fill = group)) + 
  geom_boxplot() +
  geom_point() +
  #coord_cartesian(ylim = c(1.5, 16)) + 
  stat_compare_means(comparisons = my_comparisons, label.y = c(2.5, 2.2, 2), method = "wilcox.test")+
  stat_compare_means(aes(label = paste0("\n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.55) +
  labs( # Change title, and axis names
       x = "",
       y = "Serum FITC conc. log10 transformed",
       fill = "Groups",
       caption = "Matrinem round 2 and 3") +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="lightgrey"),#Change legend box background color
    plot.caption = element_text(color = "darkgrey", size=5)) + 
  scale_fill_brewer(palette="Set1")

all_fitc_patch_log10 <- plot_all_fitc_log10 + plot_annotation (
  title = 'Intestinal permeability',
   #subtitle = 'All fitc measured',
  #tag_levels = 'A',
  caption = 'VANCO = Vancomycin, BF = Breast-Feeding, FF = Formula-Feeding\n
  FVT = Fecal Virome Transfer, SM = SM-buffer',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"),
                plot.subtitle = element_text(hjust = 0.5, face = "bold.italic"),
                plot.caption = element_text(hjust = 0.5, size=7,face = "bold")))

all_fitc_patch_log10
ggsave('../../Results/Round 4/FITC boxplot log 10.png',all_fitc_patch_log10)
```

```{r patch output}
FITC_PATCH <- all_fitc_patch + all_fitc_patch_log10
FITC_PATCH
```