### 2022-11-22 code/Histology all rounds.Rmd

---
title: "HISTO1"
author: "Kaare"
date: "6/27/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MATRINEM histo-analysis. 
AB-PAS stains quantified as % of total tissue. Experiments 2-4 collected samples for histology.

Samples are of ileum (small intestine) and colon collected from mouse pups 6 days postnatally.

```{r load packages, include=FALSE}
# Packages required
library(readxl)
library(tidyverse)
library(broom)
library(ggpubr)
library(rstatix)
library(patchwork)
```

## Load data
```{r data load}
data <- read_excel("data/raw/Histo_analysis .xlsx")

allrounds <- read_excel("data/processed/weight_development.xlsx")

```

```{r}
clean_data <- data %>% 
  mutate(ID = str_remove(ID, "-C") %>% str_remove(.,"-I") %>% str_remove(.," I")) %>% 
  mutate(ID = as.numeric(ID))

filter_joined_df <- semi_join(allrounds, clean_data, by = c("sample_id" = "ID"))

histo <- filter_joined_df %>% 
  inner_join(clean_data, by = c("sample_id" = "ID"))

histo <- histo %>% 
  pivot_longer(., `PAS-RELATIVE-TO-TOTAL(%)`:`AB-RELATIVE-TO-TOTAL(%)`, 
               names_to = "AB-PAS",
               values_to = "values")

```

```{r}
histo %>%  ### Try to wrap histo longer to have all stain values in 1 column and facet-wrap by PAS + AB
  filter(TISSUE == "ileum") %>%
  filter('AB-PAS' == "PAS-RELATIVE-TO-TOTAL(%)") %>% 
  ggplot(
    aes(
    x = group2, 
    y = values,
    fill = group2)
    ) +
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(~ exp_number,
             scales = "free") +
  labs(
    #title =  "Ileum AB-PAS staining % of total tissue", subtitle = "Faceted by experimental round (2-4)",
       y = "PAS stained as % of total tissue",
       x = "Maternal Treatment") + 
  theme_bw() +
  theme(
    axis.text.x = element_blank()
  )

```

```{r MAT vs CON}
histo %>% 
  filter(TISSUE == "colon") %>% 
   ggplot(aes(x = received_antibiotics, y = `% of total tissue stained`)) +
  geom_boxplot() + 
  facet_wrap(~ exp_number) +
  labs(title =  "COLON AB-PAS staining % of total tissue", subtitle = "Faceted by experimental round (2-4)",
       x = "Maternal Treatment") + 
  theme_bw()

```

```{r MAT vs CON}
histo %>% 
  filter(TISSUE == "colon") %>% 
   ggplot(aes(x = received_antibiotics, y = `% of total tissue stained`)) +
  geom_boxplot() + 
  facet_wrap(~ exp_number) +
  labs(title =  "COLON AB-PAS staining % of total tissue", subtitle = "Faceted by experimental round (2-4)",
       x = "Maternal Treatment") + 
  theme_bw()

```