### 2022-11-23 code/cleaning_raw_data NEW.Rmd
Updated raw data cleaning script to match project folder paths

---
title: "Raw data cleaning"
author: "Kaare D. Tranaes"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
```

#README
This is a documentation of generating the working files from the raw data files. Every step of the process is explained below.  
The raw data files can be accessed in the data/raw folder associated with this project.  
Once cleaned up, a new data file for further analysis is created and stored in the data/processed folder.

```{r load required libraries}
library(readxl) 
library(tidyverse)
library(writexl)
library(broom)
```
## Weight development data
```{r load data}
import_round1 <- read_excel("data/raw/bodyweight_round_1.xlsx"
        #                     , col_types = c("text","text", "text", "text", 
        # "text", "text", "text", "text", "numeric", 
        # "numeric", "numeric", "text")
        ) %>% 
  select(-`Pup ID`) %>% 
  rename(sample_id = `Sample ID`,
         exp_number = Round,
         received_antibiotics = `Treatment with AB`, 
         maternal_treatment = `Type of AB`,
         type_of_feed = `Feeding`,
         group = Group,
         cage = Mother,
         bodyweight_baseline = `Baseline weight`,
         bodyweight_baseline_plus_24h = `Middle of study weight`,
         bodyweight_baseline_plus_48h = `Weight End`)

import_round2 <- read_excel("data/raw/bodyweight_round_2.xlsx"
        #                     ,col_types = c("text", "text", "text", 
        # "text", "text", "text", "text", "numeric", 
        # "numeric", "numeric", "numeric","numeric", "numeric")
        ) %>% 
  rename(sample_id = `ID`,
         exp_number = Round,
         received_antibiotics = `Treatment with AntiBiotic`, 
         maternal_treatment = `Type of AB`,
         type_of_feed = `Feed`,
         group = Groups,
         cage = Cage,
         bodyweight_birthday_group_average = birthweight,
         bodyweight_birthday_plus_24H_group_average = `birth+1day`,
         bodyweight_birthday_plus_48H_group_average = `birth+2days (group mean)`,
         bodyweight_baseline = `Baseline`,
         bodyweight_baseline_plus_24h = `Study day 1`,
         bodyweight_baseline_plus_48h = `End of study`)

import_round3 <- read_excel("data/raw/bodyweight_round_3.xlsx"
        ) %>% 
  select(-id) %>% 
  rename(sample_id = `Sample ID`,
         exp_number = Exp_number,
         received_antibiotics = `Antibiotic yes/no`, 
         maternal_treatment = `Type of AB`,
         type_of_feed = `Type of feeding`,
         group = Group,
         bodyweight_baseline = `Start`,
         bodyweight_baseline_plus_24h = `Day 2`,
         bodyweight_baseline_plus_48h = `End weiight`) %>% 
  mutate(sample_id = as.numeric(sample_id),
          exp_number = as.numeric(exp_number))

import_round4 <- read_excel("data/raw/bodyweight_round_4.xlsx"
        #                     ,col_types = c("text", "text", "text", 
        # "text", "text", "text", "text", "text", 
        # "numeric", "numeric", "numeric")
        ) %>% 
  select(-mark) %>% 
  rename(sample_id = `Sample ID`, 
         group = group,
         bodyweight_baseline = `weight 0H`,
         bodyweight_baseline_plus_24h = `Weight 24H`,
         bodyweight_baseline_plus_48h = `Weight 48H`) %>% 
  mutate(cage = as.numeric(cage))

weights_all_rounds <- bind_rows(import_round1, import_round2, import_round3, import_round4) %>% 
  relocate(comments, .after = bodyweight_birthday_plus_48H_group_average) %>% 
  mutate(received_antibiotics = received_antibiotics %>% trimws(),
         maternal_treatment = maternal_treatment %>% trimws()) %>% 
  mutate(type_of_feed = if_else(type_of_feed == "Formula", "FF",
                                      if_else(type_of_feed == "Breastmilk", "BF",
                                              if_else(type_of_feed == "Breastfeeding", "BF", type_of_feed)))) %>% 
  #mutate_if(is.character,tolower) %>% 
  mutate(maternal_treatment = if_else(maternal_treatment == "Vanco", "VANCO",
                                      if_else(maternal_treatment == "Vancomycin", "VANCO",
                                      if_else(maternal_treatment == "Genta", "GENTA",
                                              if_else(maternal_treatment == "Gentamycin", "GENTA",
                                                      if_else(maternal_treatment == "Water", "CON",
                                                              if_else(maternal_treatment == "Mix", "AB",
                                                              maternal_treatment))))))) %>% 
  unite(group2, maternal_treatment:type_of_feed, sep= "-", remove = FALSE)

View(weights_all_rounds)
rio::export(weights_all_rounds, file = "data/processed/weight_development.xlsx", overwrite = TRUE)
```

## Study IDs 
```{r}
study_chars <- weights_all_rounds %>% 
  select(1:7)

rio::export(study_chars, file = "data/processed/study_id_info.xlsx", overwrite = T)

```

## Cytokines from mesoscale assays
```{r}
study_chars <- read_excel("data/processed/study_id_info.xlsx")

mesoscale_round_1 <- read_excel("data/raw/cytokines_round_1_r.xlsm")

mesoscale_round_2_and_3 <- read_excel("data/raw/cytokines_round_2_and_3.xlsx") %>% 
  rename("tissue_weight_mg" = "tissue_weight_(mg)", "exp_number" = "exp_no")

mesoscale_round_4 <- read_excel("data/raw/cytokines_round_4.xlsx") %>%
  rename("group" = "Group", "exp_number" = "exp_no")

mesoscale_all_rounds <- bind_rows(mesoscale_round_1, mesoscale_round_2_and_3, mesoscale_round_4) %>% 
  select(-group, -treatment, - exp_number)

cytokines <- left_join(study_chars, mesoscale_all_rounds, by = c("sample_id" = "sample")) %>% 
  relocate(tissue:tissue_weight_mg, .after= type_of_feed)

#rio::export(cytokines, file = "data/processed/cytokines.xlsx")
```

## Intestinal permeability - FITC-Dextran assays
```{r}
fitc_round_2_and_3 <- read_excel("data/raw/FITC_round_2-3.xlsx") %>% 
  rename(exp_number = exp_no) %>% 
  mutate("result_µg/ml_undiluted" = `result_µg/ml`*dilution_factor)

fitc_round_4 <- read_excel("data/raw/FITC_round_4.xlsx") %>% 
  mutate("result_µg/ml_undiluted" = `result_µg/ml`*dilution_factor)

fitc_all_rounds <- bind_rows(fitc_round_2_and_3, fitc_round_4) %>% 
  select(-group, - exp_number, -cage)

fitc <- left_join(study_chars, fitc_all_rounds, by = c("sample_id" = "sample")) %>%  
  filter(exp_number != "1")  # FITC-Dextran not performed in experimental round 1.

rio::export(fitc, file = "data/processed/fitc.xlsx")
```

## Gene expression - qPCR results
```{r}
### Start here. Or do it in excel.
study_chars <- read_excel("data/processed/study_id_info.xlsx")
qpcr <- read_excel("data/raw/all_genes_ct_values.xlsx")

qpcr <- qpcr %>% 
  mutate(mean_cqt = (Cq1 + Cq2)/2) %>% 
  pivot_wider(id_cols = Sample, 
              names_from = Gene,
              values_from = mean_cqt)

gene_expr <- study_chars %>%
  filter(sample_id %in% 89:112) %>% 
  left_join(., qpcr, by = c("sample_id" = "Sample"))

delta_gene_expr <- gene_expr %>%
  rename(ct_value_house_gene_HRPT = `House gene (HRPT)`) %>% 
  pivot_longer(cols = CD8:VEGFa,
               names_to = "genes",
               values_to = "ct_value") %>%
    select(sample_id, group, genes, ct_value, ct_value_house_gene_HRPT) %>% 
    filter(!is.na(ct_value)) %>%
  mutate(dct = ct_value-ct_value_house_gene_HRPT) %>% 
  group_by(genes, group) %>% 
  mutate(avg_group_dct_val = mean(dct)) %>% 
  group_by(genes) %>% 
  mutate(ddct = dct - (avg_group_dct_val[group=="CON-BF"])) %>% 
  mutate("2^-(ddct)" = 2^-ddct) %>% 
  ungroup()

# delta_gene_expr %>% 
#   group_by(genes, groups) %>% 
#   summarise(mean_fold_change = mean(relative_fold_change))

rio::export(delta_gene_expr, file = "data/processed/gene_expression_fold_changes.xlsx")

```

```{r}

delta_gene_expr %>% 
  group_by(genes, group) %>% 
  summarise(mean_fold_change = mean(`2^-(ddct)`)) %>% 
  ggplot(
    aes(x = group,
        y = mean_fold_change,
        fill = group)
  ) + 
  geom_col() +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

delta_gene_expr %>% 
  group_by(genes, group) %>% 
  ggplot(
    aes(x = group,
        y = `2^-(ddct)`,
        fill = group)
  ) + 
  geom_boxplot() +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

```

```{r}
delta_gene_expr %>%
  filter(genes %in% "CD8") %>% 
  anova(relative_fold_change ~ groups, data = .)
```