### 2022-11-26 code/Gene expression.Rmd
updated gene expression

---
title: "Gene expression"
author: "Kaare D. Tranaes"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(readxl)
library(tidyverse)
library(broom)
library(ggpubr)
library(AICcmodavg)
library(visdat)
library(rstatix)
library(gt)
library(gtsummary)
library(patchwork)
```

# Gene expression in Matrinem Experiment 4

```{r}
delta_gene_expr <- read_excel("data/processed/gene_expression_fold_changes.xlsx") %>%
   mutate(
    groups = recode_factor(  #Change the order of the groups to desired plot order
      group,
      "CON-BF"= "VANCO-BF",
      "SM-FORM" = "VANCO-SM",
      "FVT-FORM" = "VANCO-FVF"
      ))
```

## Statistical analysis of dCT levels 
```{r}
anova <- delta_gene_expr %>% 
  group_by(genes) %>% 
  anova_test(dct ~ groups) %>% 
  ungroup()

significant <- delta_gene_expr %>% 
  group_by(genes) %>% 
  do(lm(dct ~ groups,.) %>%  tidy) %>% 
  filter(term != "(Intercept)") %>% 
  filter(p.value <= 0.05)
```

## Add house gene to list
```{r}
wide_genes <- delta_gene_expr %>% 
  select(sample_id,ct_value_house_gene_HRPT, groups, genes, dct) %>% 
  pivot_wider(id_cols = c("sample_id","groups","ct_value_house_gene_HRPT"), names_from = genes, values_from = dct) %>% 
  mutate(HRPT = ct_value_house_gene_HRPT) 
```

## TAble
```{r}
Exp4_genes_table <- delta_gene_expr %>% 
  select(sample_id,ct_value_house_gene_HRPT, groups, genes, dct) %>% 
  pivot_wider(id_cols = c("sample_id","groups","ct_value_house_gene_HRPT"), 
              names_from = genes, 
              values_from = dct) %>% 
  mutate(
    HRPT = ct_value_house_gene_HRPT)  %>% 
  select(
    -sample_id, -ct_value_house_gene_HRPT) %>% 
  tbl_summary(by = groups,
              digits = everything() ~ c(1,1),
              missing = "no",
              #include = !"INFg",
              #type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ ("{mean} ({sd})")) %>%
  modify_header(
    label ~ "**Groups:**") %>% 
   #modify_footnote(stat_1 ~ "Median (IQR)", stat_2 ~ "p-value on overall group differences,Kruskall-Wallis rank sum test ") %>% 
add_p(everything() ~ "aov") %>% 
  as_gt() %>%
  tab_header(title = md("**Gene expression** *(dCT values)*")) %>%
  opt_align_table_header("center") %>% 
   tab_options(heading.title.font.size = 15,
               heading.background.color = "lightgrey",
               ) %>% 
     tab_source_note(
    source_note = "HRPT served as reference gene."
  )

Exp4_genes_table %>% 
  gtsave(path = "exports/genes/", filename = "Exp4-gene-dct-table-aov.png")
```

```{r}
delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(`2^-(ddct)`)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  ggplot(
    aes(x = groups,
        y = `2^-(ddct)`,
        fill = groups)
  ) + 
  geom_boxplot() +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

```

```{r}
my_comparisons <- list(c("VANCO-FVF", "VANCO-SM"), c("VANCO-BF", "VANCO-FVF"))
```

## Increased
```{r increased}
increased <- delta_gene_expr %>% 
 group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(`2^-(ddct)`)) %>% 
  filter(mean_fold_change >= 1.75)
```

## Increased plot 
```{r increased plot}
increased_plot <- delta_gene_expr %>% 
  filter(genes %in% increased$genes) %>% 
  mutate(across(genes, factor, levels=c("CXCL1","TNFa","MUC1","IL1b","REG3G"))) %>%
  group_by(genes, groups) %>% 
  ggplot(
    aes(x = groups,
          y = `2^-(ddct)`,
        fill = groups)
  ) +
  #geom_boxplot() +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  facet_wrap(~genes,
             nrow = 1, 
             scales = "free") + 
  labs(
       y = "Fold Δ",
       x = "",
       fill = "Groups") +
  theme_bw()+
  theme(                   
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.position = "none") + 
    scale_fill_brewer(palette="Set1") +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "VANCO-BF",
                   label.y.npc = c(0.55,0.45),
                   label.x = c(3,1))
  #stat_compare_means(comparisons = my_comparisons, label.y.npc = c(0.3,0.5), method = "wilcox.test") 
  increased_plot
```

## Decreased
```{r decreased}
decreased <- delta_gene_expr %>% 
 group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  filter(mean_fold_change < 0.6)

```

## Decreased plot
```{r decreased plot}
decreased_plot <- delta_gene_expr %>% 
  filter(genes %in% decreased$genes) %>% 
  #mutate(across(genes, factor, levels=c("MUC1","IL1b","REG3G"))) %>%
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_col() +
  facet_wrap(~genes,
             nrow = 1, 
             scales = "free") + 
  labs(
       y = "Fold Δ",
       x = "",
       fill = "Groups") +
  theme_bw()+
  theme(                   
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.position = "none") + 
    scale_fill_brewer(palette="Set1")
decreased_plot
```

## No change 
```{r no change}
no_change <- delta_gene_expr %>% 
  filter(!genes %in% c(increased$genes, decreased$genes))
```

## No change plot 
```{r plot}
no_change_plot <- no_change %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_col() +
  facet_wrap(~genes,
             nrow = 1, 
             scales = "free") +
  labs(
       y = "Fold Δ",
       x = "",
       fill = "Groups") +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="lightgrey"),
    plot.caption = element_text(color = "darkgrey", size=5)) + 
  scale_fill_brewer(palette="Set1")
```

## Patchwork
```{r}
in_de_no_patch <- (decreased_plot + increased_plot) / no_change_plot + plot_annotation(
  title = "Gene expression relative fold change",
  tag_levels = "A",
  tag_suffix = ')',
  caption = "A) Decreased relative expression.     B) Increased relative expression.     C) No change in gene expression."
) & 
  theme(plot.tag = element_text(size = 12),
        plot.caption = element_text(face = "bold"))
in_de_no_patch

```